---
title: "Fish Market Code"
author: "Yufei Yin"
linestretch: 1.5
output: pdf_document
---

# Load Package
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(leaps)
library(caret)
library(car)
```


# Load Dataset
```{r}
fish <- read_csv("fish.csv")
head(fish)
```


# Check Missing Value
```{r}
any(is.na(fish))
```


# Extra point (outlier)
```{r}
extrapoint <- tibble(Species = "Bream", Length1 = 20, Length2 = 10, Length3 = 30, Height = 15, Width = 5, Weight = 10)
fish <- bind_rows(fish, extrapoint)
```


# Data Cleaning
```{r}
fish <- fish %>% 
  filter(Weight != 0) %>%                        # Observation 41 has weight 0
  mutate(Observations = row_number()) %>% 
  mutate(Species = factor(Species)) %>% 
  rename(Vlength = Length1) %>%                  # Vertical Length
  rename(Dlength = Length2) %>%                  # Diagonal Length
  rename(Clength = Length3) %>%                  # Cross Length
  select(Observations, Species, Vlength, Dlength, Clength, Height, Width, Weight)
```

Observation 41 has weight 0 which is not reasonable, so we drop this observation.                                         
Change variable Species to factor                                                                                         
Create a new variable Observations contain the observation number.                                                                          
Length1, Length2, Length3 stand for Vertical Length, Diagonal Length, and Cross Length, So we change the variable name to Vlength, Dlength, and Clength which makes more sense.                                                                     
Finally we sort our data.


# Pair Plot
```{r}
pairs(fish[3:8])
```


# Weight vs. Vertical Length
```{r}
ggplot(data = fish, aes(x = Vlength, y = Weight, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Vertical Length", title = "Weight vs. Vertical Length")
```


# Weight vs. Diagonal Length
```{r}
ggplot(data = fish, aes(x = Dlength, y = Weight, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Diagonal Length", title = "Weight vs. Diagonal Length")
```


# Weight vs. Cross Length
```{r}
ggplot(data = fish, aes(x = Clength, y = Weight, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Cross Length", title = "Weight vs. Cross Length")
```


# Weight vs. Height
```{r}
ggplot(data = fish, aes(x = Height, y = Weight, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "Weight vs. Height")
```


# Weight vs. Width
```{r}
ggplot(data = fish, aes(x = Width, y = Weight, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "Weight vs. Width")
```


# Multicollinearity
```{r}
cor(fish[3:8])
ggplot(data = fish, aes(x = Vlength, y = Dlength, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Vertical Length", y = "Diagonal Length", title = "Diagonal Length vs. Vertical Length")
ggplot(data = fish, aes(x = Vlength, y = Clength, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Vertical Length", y = "Cross Length", title = "Cross Length vs. Vertical Length")
ggplot(data = fish, aes(x = Vlength, y = Height, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Vertical Length", title = "Height vs. Vertical Length")
ggplot(data = fish, aes(x = Vlength, y = Width, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Vertical Length", title = "Width vs. Vertical Length")
ggplot(data = fish, aes(x = Dlength, y = Clength, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Diagonal Length", y = "Cross Length", title = "Cross Length vs. Diagonal Length")
ggplot(data = fish, aes(x = Dlength, y = Height, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Diagonal Length", title = "Height vs. Diagonal Length")
ggplot(data = fish, aes(x = Dlength, y = Width, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Diagonal Length", title = "Width vs. Diagonal Length")
ggplot(data = fish, aes(x = Clength, y = Height, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Cross Length", title = "Height vs. Cross Length")
ggplot(data = fish, aes(x = Clength, y = Width, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "Cross Length", title = "Width vs. Cross Length")
ggplot(data = fish, aes(x = Height, y = Width, color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "Width vs. Height")
```


# Full Model
```{r}
full.mdl <- lm(data = fish, 
               Weight ~ Species*Vlength + Species*Dlength + Species*Clength + Species*Height + Species*Width)
summary(full.mdl)
```


# Residual Analysis
```{r, warning=FALSE, message=FALSE}
par(mfrow = c(2,2))
plot(full.mdl)
```


# Transformation
```{r}
ggplot(data = fish, aes(x = log(Vlength), y = log(Weight), color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "log Vertical Length", title = "log Weight vs. log Vertical Length")
```
```{r}
ggplot(data = fish, aes(x = log(Dlength), y = log(Weight), color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "log Diagonal Length", title = "log Weight vs. log Diagonal Length")
```
```{r}
ggplot(data = fish, aes(x = log(Clength), y = log(Weight), color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(x = "log Cross Length", title = "log Weight vs. log Cross Length")
```
```{r}
ggplot(data = fish, aes(x = log(Height), y = log(Weight), color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "log Weight vs. log Height")
```
```{r}
ggplot(data = fish, aes(x = log(Width), y = log(Weight), color = Species)) + 
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "log Weight vs. log Width")
```


# Model After Transformation
```{r}
loglog.mdl <- lm(data = fish, 
                 log(Weight) ~ Species*log(Vlength) + Species*log(Dlength) + Species*log(Clength) + 
                   Species*log(Height) + Species*log(Width))
summary(loglog.mdl)
```


# Residual Analysis After Transformation
```{r, warning=FALSE, message=FALSE}
par(mfrow = c(2,2))
plot(loglog.mdl)
```


# high leverage points
```{r, warning = FALSE, message = FALSE}
# X matrix 
attach(fish)
X <- cbind(rep(1,158), log(Vlength), log(Dlength), log(Clength), log(Height), log(Width))
# Hat matrix
H <- X %*% solve(t(X) %*% X) %*% t(X)
# hii
hii <- diag(H)
# Identify points of high Leverage
p<-ncol(X)
n<-nrow(X)
which(hii>2*p/n)
which(abs(rstandard(loglog.mdl)) >3 )
```


# outliers
```{r}
sort(cooks.distance(loglog.mdl), decreasing = TRUE)
fish_without159 <- filter(fish, Observations != 159)
loglog.mdl_without159 <-lm(data = fish_without159, 
                           log(Weight) ~ Species*log(Vlength) + Species*log(Dlength) + Species*log(Clength) +
                             Species*log(Height) + Species*log(Width))
              
coef(loglog.mdl)
coef(loglog.mdl_without159)

# Remove outliers 
fish <- filter(fish, Observations != 159)
```




# Model Selection
```{r}
loglog.mdl <- lm(data = fish, 
                 log(Weight) ~ Species*log(Vlength) + Species*log(Dlength) + Species*log(Clength) + 
                   Species*log(Height) + Species*log(Width))
step(loglog.mdl, direction = "both")
```


```{r}
loglog.mdl <- lm(data = fish, log(Weight) ~ Species + log(Clength) + log(Height) + log(Width))
summary(loglog.mdl)
```


# Further Analysis after Model Selection
```{r}
# Multicollinearity
loglog.mdl <- lm(data = fish, log(Weight) ~ Species + log(Clength) + log(Height) + log(Width))
vif(loglog.mdl)
par(mfrow = c(2,2))
plot(loglog.mdl)

# Refit Without variable Height
loglog.mdl <- lm(data = fish, log(Weight) ~ Species + log(Clength) + log(Width))
vif(loglog.mdl)
plot(loglog.mdl)

# Check high standardized residual point 14
fish_without14 <- fish[-14,]
loglog.mdl_without14 <- lm(data = fish_without14, log(Weight) ~ Species + log(Clength) + log(Width))
coef(loglog.mdl)
coef(loglog.mdl_without14)

# final model
final_mdl <- lm(data = fish, log(Weight) ~ Species + log(Clength) + log(Width))
summary(final_mdl)
```



# Cross Validation
```{r}
set.seed(10086)
for (i in 1:5){
nsamp <- ceiling(0.8*length(fish$Weight))
training_samps <- sample(c(1:length(fish$Weight)),nsamp)
train_data  <- fish[training_samps, ]
test_data <-   fish[-training_samps, ]
train.lm <- lm(data = train_data, log(Weight) ~ Species + log(Clength)  + log(Width))
summary(train.lm)
preds <- exp(predict(train.lm,test_data))
plot(test_data$Weight, preds,
     xlab = "Weight", ylab = "Predicted Weight", main = "Predicted Weight vs. Actual Weight")
abline(c(0,1))
R.sq <- R2(preds, test_data$Weight)
RMSPE <- RMSE(preds, test_data$Weight)
MAPE <- MAE(preds, test_data$Weight)
print(c(i,R.sq,RMSPE,MAPE))
}
```

```{r}
point.data <- tibble(preds, test_data$Weight)
ggplot(data = point.data, aes(x = test_data$Weight, y = preds)) + 
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(x = "Weight", y = "Predicted Weight", title = "Predicted Weight vs. Actual Weight")
```

```{r}
fish %>% group_by(Species) %>% count()
ggplot(data = fish, aes(x = fct_infreq(Species), color = Species, fill = Species)) + 
  geom_bar() +
  labs(x = "Species", title = "Species Bar Chart")
```

```{r, warning=FALSE,message=FALSE}
Whitefish <- filter(fish, Species %in% "Whitefish")
ggplot(data = Whitefish, aes(x = Width, y = Weight)) + 
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Weight vs. Width Whitefish")
```

























































